---
title: "Lab 02 - Plastic waste"
author: "Zi Li"
date: "Jan 27"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1


```{r plastic-waste-continent}
# insert code here
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```
## answer: I didn't expect this result, but after doing some research, I found Trinidad and Tobago is vulnerable to plastic pollution due to its high plastic imports and a lack of recycling infrastructure. Therefore, it's not a data error. 
## plastic waste is high in North America. 


### Exercise 2

```{r plastic-waste-density}
# insert code here
ggplot(
  data = plastic_waste,
  aes(x = plastic_waste_per_cap)
) +
  geom_density()

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent
  )
) +
  geom_density()

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density()

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.7)

```
## Ex. 2.1: Recreate the density plots above using a different (lower) alpha level that works better for displaying the density curves for all continents.
```{r}
ggplot(
  data = plastic_waste,
  aes(x = plastic_waste_per_cap)
) +
  geom_density()

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent
  )
) +
  geom_density()

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density()

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.2)

```
## EX. 2.2: aesthetic mappings like different color are linked to data variables, it's a dynamic way to represent data. Alpha level, is a characteristic of the geom that maintains visual consistency by controlling element transparency across the plot. in other word, Alpha level is the transparency level, which stay unchanged regardless of how data points are altered. 


### Exercise 3
# I've done some online research for this EX, and I filling the color and give my plot a name. 
```{r plastic-waste-violin}
# insert code here
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = continent,
    y = plastic_waste_per_cap
  )
) +
 geom_boxplot(fill = "skyblue") +
  ggtitle("Box Plot: Plastic Waste per Capita by Continent")

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = continent,
    y = plastic_waste_per_cap
  )
) +
geom_violin(fill = "lightpink", alpha = 0.2) +
  ggtitle("Violin Plot: Plastic Waste per Capita by Continent")
```

## answer for Ex. 3.1: violin plot reveal the shape and density of the data; and box plot reveal the median and it also allow for comparisons of the central tendency.

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-mismanaged}
# insert code here
```

```{r plastic-waste-mismanaged-continent}
# insert code here
```

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r recreate-viz}
# insert code here
```


